---
title: "reverse_engineering"
output: html_document
date: "2025-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dplyr)
library(stringr)
```

**Code Run**

```{r}
data <- read.csv("data/updated_feb2024_bln_reuters_climate_finance_2015to2020.csv")

```

**Question 1: “Across the globe, donors reported more than $25 billion in funding they claim is linked to renewable energy. At least another $5.6 billion went to projects they claim would help countries prepare for or respond to climate-related disasters. Many of their reports contain too little detail to verify these claims, Reuters found.”**

**Answer 1: There are multiple sources of renewable energy that could suggest $25 billion is linked to renewable energy. For example, mitigation, cross-cutting, and crosscutting. There were multiple names for the same types of support so just by filter those key names, we were able to come up with an outcome. This is effective because it brings fourth multiple ways countries organize their findings by certain keywords. Not every country will spell crosscutting with a hyphen but some do. Also, the data was given in scientific notation which isn't quite helpful so turning it into real numerical value was helpful.**

```{r}
renewable_energy <- data |>
  group_by(type_of_support) |>
  summarise(total_funding = sum(corrected_usd, na.rm = TRUE)) |>
  arrange(desc(total_funding))

renewable_energy |>
  mutate(total_funding = format(total_funding, big.mark = ",", scientific = FALSE))
```




