---
title: "reverse_engineering"
output: html_document
date: "2025-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dplyr)
library(stringr)
library(janitor)
```

**Code Run**

```{r}
data <- read.csv("data/updated_feb2024_bln_reuters_climate_finance_2015to2020.csv")

data |>
  clean_names()


```

  **Data explained What is each row? Each row is a different id or characteristic like year, recipient country, money in usd, all attributes of each instance of a wealthy nation "spending money" on climate related efforts. Where did it come from? The data comes from the Reuters-Stanford Climate Finance Project catalog published by Reuters originally but open for public domain.Does it have any known errors? It has a couple errors that make data have N/A until the names are cleaned under certain categories like certain foreign currencies. What does that mean about the caveat? That there is no way to know for 100% certainty what all the data actually is and what numbers they truly are to pin point accuracy. There's no way to know for certain.**



**Question 1: From 2015 to 2020, 35 governments reported a total of more than $182 billion in grants, loans, bonds, equity investments and other contributions.”**

**Answer 1:I wanted to see from the 35 governments between 2015 to 2020 how much money was spent toward loans, bonds, equity investments, and other contributions. I created a renamed the corrected_usd to total_sum and rearranged by descending.**

```{r}
governments <- data |>
  group_by(reporting_party) |>
  summarise(total_sum = sum(corrected_usd)) |>
  arrange(desc(total_sum))

governments |>
  filter(reporting_party != "Monaco") |>
  summarise(total_sum_excluding_monaco = sum(total_sum))


```


**Question 2: Developed nations reported more than 40,000 direct contributions toward the finance target, totaling more than $182 billion, from 2015 to 2020, the last year for which data is available.**

**Answer 2: The data starts in 2015 and goes through 2020 and equals about $189,370,929,494 with multiple direct contributions toward the finance target such as funding_source**

```{r}
direct_contributions <- data |>
  filter(year >= 2015 & year <= 2020) |>
  summarise(count = n(),total = sum(corrected_usd, na.rm = TRUE))
```





**Question 3: Loans totaling $267.5 million for upgrades to a metro system in Mexico**

**Answer 3:**

```{r}
Mexico <- data |>
  filter(if_any(everything(), ~ str_detect(tolower(as.character(.x)), "loan")))

Mexico |>
  filter(if_any(everything(), ~ str_detect(tolower(as.character(.x)), "metro")))
  
Mexico |>
  filter(str_detect(tolower(recipient), "mexico"))|>
  summarise(count = n(),total = sum(corrected_usd, na.rm = TRUE))
```
