---
title: "reverse_engineering"
output: html_document
date: "2025-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dplyr)
library(stringr)
library(janitor)

options(scipen=999)

```

**Code Run**

```{r}
data <- read.csv("data/updated_feb2024_bln_reuters_climate_finance_2015to2020.csv")

data |>
  clean_names()


```

  **Data explained What is each row? Each row is a different id or characteristic like year, recipient country, money in usd, all attributes of each instance of a wealthy nation "spending money" on climate related efforts. Where did it come from? The data comes from the Reuters-Stanford Climate Finance Project catalog published by Reuters originally but open for public domain.Does it have any known errors? It has a couple errors that make data have N/A until the names are cleaned under certain categories like certain foreign currencies. What does that mean about the caveat? That there is no way to know for 100% certainty what all the data actually is and what numbers they truly are to pin point accuracy. There's no way to know for certain.**



**Question 1: From 2015 to 2020, 35 governments reported a total of more than $182 billion in grants, loans, bonds, equity investments and other contributions.”**

**Answer 1:I wanted to see from the 35 governments between 2015 to 2020 how much money was spent toward loans, bonds, equity investments, and other contributions. I originally filtered by financial_instruments and searched for  "grant", "loan", "equity investment". I also, grouped by reporting_party and total_sum = sum(corrected_usd) and arranged them by descending sums. This didn't necessarily work because I wasn't finding the specific financial instrument that was 7,000,000,000 from the total $189 billion of all spending. I found "other" that had about $7 billion so I filtered by that financial instrument. **

```{r}
governments <- data |>
  filter(!is.na(reporting_party)) |>
  filter(!is.na(corrected_usd)) |>
 filter(financial_instrument != "Other") |>
 summarise(total_sum = sum(corrected_usd)) 

governments |>
 group_by(financial_instrument) |>
  summarise(total = sum(corrected_usd, na.rm = TRUE)) |>
  arrange(desc(total))
```


**Question 2: The United States agreed to lend $19.5 million to developers of a Marriott hotel franchise in Cap-Haitien, Haiti.**

**Answer 2: **

```{r}
hotel <- data |>
  filter(reporting_party == "united states") |>
  filter(recipient == "haiti") |>
  filter(if_any(everything(), ~ str_detect(tolower(as.character(.x)), "marriott"))) |>
  summarise(total_sum = american_dollar) 
  
```





**Question 3: Loans totaling $267.5 million for upgrades to a metro system in Mexico**

**Answer 3:**

```{r}
Mexico <- data |>
  filter(if_any(everything(), ~ str_detect(tolower(as.character(.x)), "loan")))

Mexico |>
  filter(if_any(everything(), ~ str_detect(tolower(as.character(.x)), "metro")))
  
Mexico |>
  filter(str_detect(tolower(recipient), "mexico"))|>
  summarise(count = n(),total = sum(corrected_usd, na.rm = TRUE))

```
