---
title: "data_analysis_group4"
authors: "Razak Diallo, Lea Glaser, Deja Jones, Alexander Taylor"
date: "2025-12-06"
output: html_document
---

# Load Libraries
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidycensus)
```


# Load and clean data
```{r}


state_payments <- read_csv("data/state_payments.csv") |>
  clean_names() |>
  mutate(
    date = ymd_hms(date),
    amount = str_replace_all(amount, "\\$", ""),
    amount = str_replace_all(amount, "\\,", ""),
    amount = as.numeric(amount)
         )

glimpse(state_payments)


```
# Question 1: What is the median income of the zip codes where the top five vendors are located?
# Findings: All five are displayed below. Two of the zip codes have a median income greater than $120,000, one is just under $100,000 and the other, which has the highest amount of payments, has a median income of about $60,000. I would want to look more the last zip code to find because it seems to be an outlier: it has a lower median income but a high amount of payments. 

# NOTE: There doesn't seem to be median income data for 10041. How do you suggest I get around this?
```{r}
# Find zip codes with the top five vendors

state_payments |>
  group_by(date) |>
  summarise(count = n())

q1_top5 <- state_payments |>
  group_by(vendor_zip) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount)) |>
  filter(row_number() <= 5)

# Get median income data
#q1_vars <- load_variables(2023, "acs5", cache = TRUE)
#B19013_001


q1_vars <- load_variables(2023, "acs5", cache = TRUE)

q1_med_income <- get_acs(geography = "zcta", variables = "B19013_001", year = 2023) |>
  rename(
    vendor_zip = GEOID,
    med_income = estimate
  ) |>
  select(vendor_zip, med_income)

# Join dataframes

q1_top5 <- q1_top5 |>
  left_join(q1_med_income)

q1_top5


```

# Question 2: What year did the State of Maryland spend the most? How about the least?
# Findings: The State of Maryland spent the most in 2025, spending $43,703,480,470 and the least was in 2008 (if not counting 2026) where it spent $19,879,214,691
```{r}
years <- state_payments |>
  group_by(fiscal_year) |>
  summarise(total_spending = sum(amount, na.rm = TRUE)) |>
  arrange(desc(total_spending))

```

# Question 3: What is the total amount of payments in each zip code?
# Findings: 
```{r}

```

# Question 4: How many payments were made each month?
# Findings: 
```{r}

```

# Question 5: What agency spent the most money in each county?
# Findings: 
```{r}

```
